Case for Support

### 2021-11-17 application/Je-S_sections.tex

1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.

2. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment to implement:
2a. video-based  behavioural analysis
2b. real-time interfacing of population-scale activity with laboratory control  
2c. evaluate competing data analytic approaches within a single experimental framework

3. Enhance the quality and reproducibility of systems neuroscience through:
3a. fully open-source and public-facing development
3b. rapid release of, and community engagement around, new tools
3c. provision of standard implementations of analysis and control methods, with the ability to distribute automated
descriptions of analysis pipelines for reproducibility.

4. Extend and support the Bonsai user community through code and forum engagement and development workshops.

@@

Here we propose to provide the neuroscience community a software tool that will dramatically increase the efficiency of experimental control and data analysis. We will do so by developing a new set of functionalities to an existing software platform, Bonsai. Bonsai is a fully integrated software environment that emphasises performance, flexibility, and ease-of-use, allowing scientists with no previous programming experience to quickly develop their own high-performance data acquisition and experimental control systems.

@@

Resource is Bonsai.
Monitor animal welfare

>>>>

1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment. \textcolor{red}{I believe this is already done. All existing bonsai community packages already depend on the existence of the extendable plugin architecture -- perhaps you mean "languages" rather than "packages"? i.e. MATLAB, Python, R, in that case I would rewrite as below}

1. Develop, test and deploy new foreign language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.

2. Define a standard API and set of operators appropriate for composing machine-learning functionality into a Bonsai workflow, specifically to enable:
2a. learning
2b. inference
2c. visualisation
2d. model-comparison and validation
2e. exchange of models and hyper-parameters

3. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment using the standard API specified above, to implement:
3a. video-based behavioural analysis
3b. real-time interfacing of population-scale extracellular activity with laboratory control  
3c. evaluation of competing data analytic approaches within a single experimental framework

4. Enhance the quality and reproducibility of systems neuroscience through:
4a. fully open-source and public-facing development
4b. rapid release of, and community engagement around, new tools
4c. provision of standard implementations of analysis and control methods, with the ability to distribute fully automated and reproducible analysis pipelines.

5. Extend and support the Bonsai user community through code and forum engagement and development workshops.

@@

Here we propose to provide the neuroscience community with a software tool that will dramatically increase the efficiency of experimental control and data analysis. We will do so by developing a new set of functionalities to an existing software platform, Bonsai. Bonsai is a fully integrated software environment that emphasises performance, flexibility, and ease-of-use, allowing scientists with no previous programming experience to quickly develop their own high-performance data acquisition and experimental control systems.

@@

The main beneficiaries of this proposal will be experimental labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments. While more and more automated Machine Learning platforms are becoming generally available across many data science fields, they either focus on making available completely general-purpose tools which require deep expertise and training to construct and train new algorithms and models from scratch, or they encapsulate some pre-trained and domain-specific model which is useful only for very specific classes of problems (e.g. face recognition, pose estimation).

A middle ground where existing algorithms can be easily composed by non-experts in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions is currently entirely absent. In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for new bespoke algorithms and analysis procedures is constant and critical. If successful, this proposal will entirely transform the landscape of Machine Intelligence applied to experimental lab research in biology, by empowering non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.

In addition to the researchers themselves, the standard analysis pipelines will allow everyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data. As an example, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions. They use Bonsai as it allows them to prototype and design entirely new monitoring systems on-site, without the need to engage an expensive software development process to answer simple questions. With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment are entirely unthinkable, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.
